---
title: "CARDIAC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CARDIAC STEP COUNT PLOTS


```{r}
library('rio') # import
library('janitor') # cleaning column names
library('ggplot2') # visualization
library('tidyverse') # data management
library('dplyr')  # data management
library(plotly)
library(ggthemes)
library(rstatix)

cr <- import('CardiacRehabilitatio-Functionalcapacitype_DATA_LABELS_2024-02-08_1528 (1).csv')

cr <- clean_names(cr)
names(cr)
```

# Objective
# 1. Compare Week 1's target, what the device recorded and what the participant reported.
# 2. Compare trend of subsequent weeks


# Data Cleaning and Transformation

# First obtain the participants asked to exercise
# get rows 1:93, those who were asked to exercise
##exercise <- cr[1:93, ]

```{r}
exercise <- cr[1:93, ]
exercise_hbcr <- exercise[grepl("HBCR", exercise$event_name), ]

# WEEK 1 STEP COUNT COMPARISON
week_1 <- exercise_hbcr[c("wk_1_step_count_target","wk_1_step_count_achieved_device")]

# merge study IDs to steps device+
id <- exercise_hbcr[,1:2]
# bind week 1 with study id
week_1_df <- cbind(id, week_1)

## dropping study ID 5, value appears to be an outlier
week_1_df <- week_1_df[-1, ]
rownames(week_1_df) <- NULL
```

## Plots

```{r pressure, echo=FALSE}
## Plot week 1 comparison: Target vs Achieved
# Pivoted the data frame to a long format
long_week_1 <- week_1_df %>% 
  pivot_longer(
    cols = c("wk_1_step_count_target","wk_1_step_count_achieved_device"),
    names_to = 'study_weeks',
    values_to = 'steps_taken'
  )
# change study_id column to character
# So that y axis isn't a continuous variable
long_week_1$study_id <- as.character(long_week_1$study_id)

# visualize the plot
long_week_1 %>% 
  ggplot(aes(x=steps_taken, y=study_id, group= study_id, color=study_weeks))+
  geom_line()+
  geom_point()+
  labs(x = "Step_Counts", y = "Study IDs", title = "Week 1 Step Count Target vs Achieved")+
  theme_minimal()

## SD $ Variance
week_1_df$difference <- week_1_df$wk_1_step_count_target - week_1_df$wk_1_step_count_achieved_device
sd_week_1 <- sd(week_1_df$difference, na.rm = TRUE)
var_week_1 <- var(week_1_df$difference, na.rm = TRUE)
print(sd_week_1)

```
## Those told not to monitor
```{r}

exercise_not <- cr[94:225, ] # those who were told not to monitor steps
# 2nd, obtain the Arm: Home Based row
exercise_not_hbcr <- exercise_not[grepl("HBCR", exercise_not$event_name), ]

# column names
names(exercise_not_hbcr)
# not to monitor
e_not_week_1 <- exercise_not_hbcr[c("wk_1_step_count_achieved_participant_report_via_phone","step_count_achieved_participant_report_via_phone")]


# merge study IDs to steps device
id <- exercise_not_hbcr[,1:2]
# bind week 1 with study id
e_not_week_1_df <- cbind(id, e_not_week_1)

## dropping study ID 5, value appears to be an outlier
# week_1_df <- week_1_df[-1, ]
rownames(e_not_week_1_df) <- NULL
```

```{r pressure, echo=FALSE}
## Plot week 1 comparison vs followup
# Pivoted the data frame to a long format
e_not_long_week_1 <- e_not_week_1_df %>% 
  pivot_longer(
    cols = c("wk_1_step_count_achieved_participant_report_via_phone","step_count_achieved_participant_report_via_phone"),
    names_to = 'study_weeks',
    values_to = 'steps_taken'
  )


# change study_id column to character
# So that y axis isn't a continuous variable
e_not_long_week_1$study_id <- as.character(e_not_long_week_1$study_id)

# visualize the plot
e_not_long_week_1 %>% 
  ggplot(aes(x=steps_taken, y=study_id, group= study_id, color=study_weeks))+
  geom_line()+
  geom_point()+
  labs(x = "Step_Counts", y = "Study IDs", title = "Week 1 Step Count  vs Week 2")+
  theme_minimal()


names(e_not_week_1_df)
## SD $ Variance
e_not_week_1_df$difference <- e_not_week_1_df$wk_1_step_count_achieved_participant_report_via_phone - e_not_week_1_df$step_count_achieved_participant_report_via_phone

sd_week_1 <- sd(e_not_week_1_df$difference, na.rm = TRUE)
var_week_1 <- var(week_1_df$difference, na.rm = TRUE)
print(sd_week_1)

```

# Objective 2

```{r}
# Trends Over the weeks
steps <- cr[, grepl("Step count", names(cr))]
steps_device <- cr[, grepl("device", names(cr))]

# merge study IDs to steps device
study_id <- cr[,1:2]

## clean column names
names(steps_device) <- NULL
# New column names
names <- c("1","2","3","4","5","6","7","8","9","10","11","12")
# Assign new column names
names(steps_device) <- names


study_steps_device <- cbind( study_id, steps_device)
names(study_steps_device)



## those told to monitor
# get rows 1:93, those who were asked to exercise
e_study_steps_device <- study_steps_device[1:93, ]


# get HBCR row
study_hbcr <- e_study_steps_device[grepl("HBCR", e_study_steps_device$event_name), ]

# drop 2nd column
study_hbcr <- study_hbcr[,-2]
rownames(study_hbcr) <- NULL

names(study_hbcr)
```

# visualize
```{r}
### Visualization
study_hbcr$study_id <- as.character(study_hbcr$study_id)
long_hbcr <- study_hbcr %>%
  pivot_longer(
    cols = 2:13,
    names_to = "weeks",
    values_to = "step_taken"
  )

# change weeks to time format
# -change to numeric

class(long_hbcr$weeks)
long_hbcr$weeks <- as.numeric(long_hbcr$weeks)

long_hbcr$study_id <- as.character(long_hbcr$study_id)



reference_date <- as.Date(("2024-01-01"), by = "week", length.out = 12)
# long_hbcr$weeks <- as.Date( reference_date + (long_hbcr$weeks *7))

#long_hbcr$weeks <- as.Date(long_hbcr$weeks)


# Function to create week labels
week_labels <- function(dates) {
  week_numbers <- seq_along(dates)
  paste("weeks", week_numbers)
}

# line plot
long_hbcr_poltly <- long_hbcr %>% 
  ggplot(aes(x=weeks, y= step_taken, color=study_id))+
  geom_point()+
  geom_line()+
  labs(title="Trends of steps taken by those told to Monitor")+
  scale_x_continuous(labels = scales::number_format(accuracy = 1))+
  theme_few()+
  theme(legend.position = "none")
plot_ids <- ggplotly(long_hbcr_poltly)
plot_ids

get_summary_stats(study_hbcr)
```

#Step count target
```{r}
steps_target <- cr[, grepl("step_count_target", names(cr))]

# merge study IDs to steps device
study_id <- cr[,1:2]

## clean column names
names(steps_target) <- NULL
# New column names
names <- c("1","2","3","4","5","6","7","8","9","10","11","12")
# Assign new column names
names(steps_target) <- names


study_steps_target <- cbind( study_id, steps_target)
names(study_steps_target)



## those told to monitor
# get rows 1:93, those who were asked to exercise
e_study_steps_target <- study_steps_target[1:93, ]


# get HBCR row
target_study_hbcr <- e_study_steps_target[grepl("HBCR", e_study_steps_target$event_name), ]

# drop 2nd column
target_study_hbcr <- target_study_hbcr[,-2]
rownames(target_study_hbcr) <- NULL

names(study_hbcr)
```

```{r}
### Visualization
target_study_hbcr$study_id <- as.character(target_study_hbcr$study_id)
target_long_hbcr <- target_study_hbcr %>%
  pivot_longer(
    cols = 2:13,
    names_to = "weeks",
    values_to = "step_taken"
  )

# change weeks to time format
# -change to numeric

class(target_long_hbcr$weeks)
target_long_hbcr$weeks <- as.numeric(target_long_hbcr$weeks)

long_hbcr$study_id <- as.character(long_hbcr$study_id)



# Function to create week labels
week_labels <- function(dates) {
  week_numbers <- seq_along(dates)
  paste("weeks", week_numbers)
}

# line plot
target_long_hbcr_poltly <- target_long_hbcr %>% 
  ggplot(aes(x=weeks, y= step_taken, color=study_id))+
  geom_point()+
  geom_line()+
  labs(title="Trends of Target steps by those told to Monitor")+
  scale_x_continuous(labels = scales::number_format(accuracy = 1))+
  theme_few()+
  theme(legend.position = "none")
plot_ids <- ggplotly(target_long_hbcr_poltly)
plot_ids

get_summary_stats(study_hbcr)
```



```{r}
value1 <- target_long_hbcr$step_taken
value2 <- not_e_long_hbcr$step_taken

result <- t.test(value1, value2, paired = TRUE)
print(result)
# Print detailed results
cat("t-statistic:", result$statistic, "\n")
cat("Degrees of freedom:", result$parameter, "\n")
cat("p-value:", result$p.value, "\n")
cat("Confidence interval:", result$conf.int, "\n")
cat("Mean of the differences:", result$estimate, "\n")

```






```{r}

not_exercise <- cr[94:225, ] # those who were told not to monitor steps
# 2nd, obtain the Arm: Home Based row
not_exercise_hbcr <- not_exercise[grepl("HBCR", not_exercise$event_name), ]
# 

steps_phone <- not_exercise_hbcr[, grepl("phone", names(not_exercise_hbcr))]

# # merge study IDs to steps phone
not_e_id <- not_exercise_hbcr[,1:2]

# # bind study id with steps phone
not_e_study_steps_phone <- cbind( not_e_id, steps_phone)

rownames(not_e_study_steps_phone) <- NULL
# drop 2nd column
not_e_study_steps_phone <- not_e_study_steps_phone[,-2]

##
# New column names
not_e_names <- c("study_id","1","2","3","4","5","6","7","8","9","10","11","12")
# Assign new column names
names(not_e_study_steps_phone) <- not_e_names


```

```{r}
### Visualization
# those told not to exercise
not_e_study_steps_phone$study_id <- as.character(not_e_study_steps_phone$study_id)

## getting 31 obs
num_rows <-nrow(not_e_study_steps_phone)
not_e_study_steps_phone <- not_e_study_steps_phone[1:(num_rows-13), ]

not_e_long_hbcr <- not_e_study_steps_phone %>%
  pivot_longer(
    cols = 2:13,
    names_to = "weeks",
    values_to = "step_taken"
  )

# change weeks to time format
# -change to numeric

class(not_e_long_hbcr$weeks)
not_e_long_hbcr$weeks <- as.numeric(not_e_long_hbcr$weeks)

not_e_long_hbcr$study_id <- as.character(not_e_long_hbcr$study_id)



reference_date <- as.Date(("2024-01-01"), by = "week", length.out = 12)
#not_e_long_hbcr$weeks <- as.Date( reference_date + (not_e_long_hbcr$weeks *7))




# line plot
not_e_long_hbcr_poltly <- not_e_long_hbcr %>% 
  ggplot(aes(x=weeks, y= step_taken, color=study_id))+
  geom_point()+
  geom_line()+
  labs(title="Trends of steps taken by those told NOT to Monitor")+
  scale_x_continuous(labels = scales::number_format(accuracy = 1))+
  theme_few()+
  theme(legend.position = "none")

not_e_plot_ids <- ggplotly(not_e_long_hbcr_poltly)
not_e_plot_ids

get_summary_stats(not_e_study_steps_phone)
summary(not_e_study_steps_phone)
```


